---
title: "p8105_hw1_td2749"
author: "Tvisha R. Devavarapu"
date: "2022-09-19"
output: github_document
---

**Loading the required packages:**
```{r Loading required packages, message = FALSE}
library(tidyverse)
library(ggplot2)
```

## Problem 1

```{r P1_installing_dataset, eval = FALSE}
install.packages("palmerpenguins")
```

Loading the `penguins` dataset from the `palmerpenguins` package: 
```{r P1_loading_dataset}
data("penguins", package = "palmerpenguins")
```

```{r P1_calculating_values, include = FALSE}
# Data in the dataset, including names / values of important variables
colnames(penguins)

# The size of the dataset (using nrow and ncol)
nrow(penguins)
ncol(penguins)
```
```{r P1_dropping_na_rows_flipper_length, echo = FALSE}
# Creating a new df with flipper length NA value rows deleted. 
nonNA_flipperlength_df = drop_na(penguins, flipper_length_mm)

# The mean flipper length
mean_flipper_length = mean(nonNA_flipperlength_df$flipper_length_mm)
```

_**Dataset Description**_

The given dataset consists of information regarding penguins from the years 2007-2009. It holds **`r nrow(penguins)` rows and `r ncol(penguins)` columns**. The columns contain data pertaining to _'`r colnames(penguins)`'_. Data recordings for bill length, bill depth, and flipper length are made in millimeters (mm). body_mass is recorded in grams (g). The species include _'`r unique(penguins$species)`'_ from the _'`r unique(penguins$island)`'_ islands. The mean flipper length recorded for the listed penguins is **`r round(mean(nonNA_flipperlength_df$flipper_length_mm), digits = 2)` mm** (millimeters). 

NOTE: 2 rows from the `penguin` dataframe with NA values in the `flipper_length_mm` column were omitted to perform the above mean calculation.

_**Filpper Length Vs. Bill Length: Visualization Via Scatterplot**_

```{r P1_scatterplot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave('bill_vs_flipper_scatter.png')
```


## Problem 2

```{r P2_creating_df}
df_2 = 
  tibble(
    norm_vec = rnorm(10),
    log_vec = norm_vec > 0,
    char_vec = c("I", "love", "listening", "to", "Seventeen", "NCT", "Treasure", "and", "BTS", "!"),
    fac_vec = factor(c("NYC", "SD", "LA", "LA", "SD", "SD", "SD", "NYC", "NYC", "NYC"))
  )
```

_**Taking means of each variable**_
```{r P2_taking_means}
#Mean of numerical values
mean(pull(df_2, norm_vec))

#Mean of logical values
mean(pull(df_2, log_vec))

#Mean of character values
mean(pull(df_2, char_vec))

#Mean of vector values
mean(pull(df_2, fac_vec))
```

Based on the above results, it is possible to take means of numeric and logical values but not of character or factor values. 

```{r P2_as_numeric, eval = FALSE}
log_num = as.numeric(df_2$log_vec)
char_num = as.numeric(df_2$char_vec)
fac_num = as.numeric(df_2$fac_vec)
```

Calling `as.numeric` on the logical values converts them into zeroes (FALSE) and ones (TRUE). This explains why it is possible to run the `mean` function on the logical vector. Character values on the other hand don't seem to be getting converted into numeric values. Calling `as.numeric` on them throws up a 'Warning Message' of _"NAs introduced by coercion"_. Character values not being able to be converted to numbers explains why it is not possible to call the `mean` function on a character vector. Upon being called with `as.numeric`, factor values are getting converted into numeric representations of the number of classes in the vector in context. As the numeric representation only seems to be functioning as some sort of identification aspect rather than actually converting the factor values into numbers, it makes sense that there is no possibility of calling their mean. From a reasoning perspective, calling the mean of a logical vector gives an indication of the proportion of TRUE and FALSE values present in it (as they are only 0s and 1s so a 0.7 mean recorded from a collection of 10 values indicates the presence of 7 TRUEs). However, there is not much that can be inferred from the mean calculated from the as.numeric converted factor values as they would widely differ based on the number of classes and values present and offer no specific information regarding the factor vector in reference. 
