---
title: "p8105_hw1_td2749"
author: "Tvisha R. Devavarapu"
date: "2022-09-19"
output: github_document
---

Loading the required packages:
```{r Loading required packages}
library(tidyverse)
library(ggplot2)
```

## Problem 1

```{r P1_installing_dataset, eval = FALSE}
install.packages("palmerpenguins")
```

Loading the `penguins` dataset from the `palmerpenguins` package: 
```{r P1_loading_dataset}
data("penguins", package = "palmerpenguins")
```

```{r P1_calculating_values, include = FALSE}
# Data in the dataset, including names / values of important variables
colnames(penguins)

### NOTE TO SELF: Make a detailed description of the types, areas, means, etc. 

# The size of the dataset (using nrow and ncol)
nrow(penguins)
ncol(penguins)
dim(penguins)
length(penguins)
```
```{r P1_dropping_na_rows_flipper_length, echo = FALSE}
# Creating a new df with flipper length NA value rows deleted. 
nonNA_flipperlength_df = drop_na(penguins, flipper_length_mm)

# The mean flipper length
mean_flipper_length = mean(nonNA_flipperlength_df$flipper_length_mm)
```

The given dataset consists of information regarding penguins from the years 2007-2009. It holds **`r nrow(penguins)` rows and `r ncol(penguins)` columns**. The columns contain data pertaining to `r colnames(penguins)`. 
The mean flipper length recorded is `r mean(nonNA_flipperlength_df$flipper_length_mm)` mm (millimeters).

NOTE: 2 rows from the `penguin` dataframe with NA values in the flipperlength column were omitted to perform the mean calculation. 

```{r P1_scatterplot}
# Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable 

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave('P1_scatterplot-1.png')
```



`r `
